Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-09-06 Latex_files/Thesis/chapters/chapter 2/main.tex
updates 06/09/2024

The scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the both the scattering force, and the thermal 
energy of the particle. Describing and calculating the optical forces is 
dependent on the ratio between the particle size and the trapping wavelength.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

@@

Vigilante et al compiled together a python package that combines both \textit{ott} 
and \textit{MSTM} to simulate the behaviour of spherical aggregates within 
a predefined optical trap \cite{Vigilante2020}.

\subsection{Simulation parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation state of 
the particle, orientations are reported as quaternions which can be readily 
transformed into rotation matrices. The matrix describes the direction of the 
particles 3 orientation axis $U_x$, $U_y$, and $U_z$; the latter is oriented 
so that it is collinear with the centres of the two spheres. The origin of 
the 3 axis' is set on the particle's centre of mass, this is also used to 
define it's displacement from the origin of the laboratory frame.

Furthermore we expand upon their work to create a fully functional quadrant 
photo-diode to simulate the response from a calibration test. This builds 
upon the work from \cite{Rohrbach2002} which applied the fundamental 
principles of Lorenz-Mie theory to replicate the response signal of a QPD 
being used in back focal-plane interferometry.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulated Quadrant Photodiode}
\label{sec:simulated_QPD}
In order to simulate a typical experimental set up with a QPD installed 
as a position detection system we need to evaluate the total magnitude of 
the electric field incident on the photo-diode surface. While trapping a 
micro-particle, the scattered and incident fields combine together and 
interfere with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD surface, the 
total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}

>>>>

The scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the both the scattering force, and the thermal 
energy of the particle.

When a sphere is located at the centre of a laser the gradient force will
try to recentre the sphere if it is displaced. If the restoring force is
proportional to the sphere's displacement this is referred to as a harmonic
trap. The ratio of the restoring force to the displacement is called the 
trap stiffness (denoted by $\kappa$), and is an important characteristic 
of the trap. Shown below is a typical force displacement curve for a single
sphere in a focused Gaussian beam.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.78\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both the laser
power and the contrast in refractive indices between the medium and the
particle. Fully describing and calculating the optical forces is dependent on 
the ratio between the particle size and the trapping wavelength. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

@@

Vigilante et al compiled together a python package that combines both 
\textit{ott} and \textit{mstm} to simulate the behaviour of spherical 
aggregates within a predefined optical trap \cite{Vigilante2020}. Now 
\textit{ott} does have inbuilt $T$-matrix modules that can be used to 
model the scattering from different shaped particles; however, for 
aggregates of spheres the resultant force is nonsensical (see Chapter 
4 for further details). As such we rely on the \textit{mstm} for 
simulating dimers.

\subsection{Simulation parameters}
\label{sec:sim_parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation state of 
the particle, orientations are reported as quaternions which can be readily 
transformed into rotation matrices. The matrix describes the direction of the 
particles 3 orientation axis $U_x$, $U_y$, and $U_z$. The latter is oriented 
so that it is collinear with the centres of the two spheres; the direction of
vector $U_z$ is such that it goes from the centre of $a_{II}$ to the centre of
$a_{I}$. The origin of the 3 axis' is set on the particle's centre of mass, 
this is also used to define it's displacement from the origin of the laboratory 
frame. The diffusion tensor of the dimer is more complex than a sphere's as we 
must consider the rotational effects experienced by the dimer. We use the 
analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided the 
solution coefficients depending on the dimer's size ratio ($a_I/a_{II}$). Rather 
than recalculate the coefficients for every possible configuration of dimer we 
use a spline fitting function to approximate the diffusion tensor. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulated Quadrant Photodiode}
\label{sec:simulated_QPD}

Beyond merely simulating the forces experienced by a spherical aggregate we
also developed a simulative quadrant photo diode to replicate the results
from a typical calibration test. This builds upon the work from \cite{Rohrbach2002} which applied Lorenz-Mie theory to replicate the response signal of a QPD 
being used in back focal-plane interferometry. However, rather than be 
constrained to individual spheres we can now consider the expected response 
from any type of spherical aggregate.

In order to simulate a typical experimental set up with a QPD installed 
as a position detection system we need to evaluate the total magnitude of 
the electric field incident on the photo-diode surface. While trapping a 
micro-particle, the scattered and incident fields combine together and 
interfere with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD surface, the 
total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}
